---
title: 'Reproducible Research: Project 1'
output: 
  html_document: 
    fig_caption: yes
---
## Reading in the dataset and processing the data
```{r EPL, echo=TRUE}
library(dplyr)
library(lubridate)
library(ggplot2)
library(tidyr)

activity <- read.csv("data/activity.csv") %>%
  mutate(date=ymd(date))
```

## Histogram of the total number of steps taken each day
```{r hist1, echo=TRUE}
activity %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarise(total_steps=sum(steps, na.rm=T)) %>%
  ggplot(aes(x = total_steps)) +
        geom_histogram(binwidth = 1000) +
        labs(title = "Daily Steps", x = "Total Steps", y = "Frequency")
```

## Mean and median number of steps taken each day
```{r mean and median, echo=TRUE}
activity %>%
  group_by(date) %>%
  summarise(mean=mean(steps, na.rm=T), median=median(steps, na.rm=T))
```

## Time series plot of the average number of steps taken
```{r time series plot, echo=TRUE}
activity %>%
  group_by(date) %>%
  summarise(mean=mean(steps, na.rm=T), median=median(steps, na.rm=T)) %>%
  ggplot(aes(x=date, y=mean)) +
  geom_line() 
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r best interval, echo=TRUE}
activity %>%
  group_by(interval) %>%
  summarise(mean=mean(steps, na.rm=T)) %>%
  ggplot(aes(x=interval, y=mean)) +
  geom_line()

activity %>%
  group_by(interval) %>%
  summarise(mean=mean(steps, na.rm=T)) %>%
  top_n(1, wt=mean)
```

## Strategy for imputing missing data
#### How many missing values in each column ?
```{r total missing values, echo=TRUE}
table(is.na(activity$steps))
table(is.na(activity$date))
table(is.na(activity$interval))
```
There are missing values only in the "steps" column.

#### Imputing missing values
```{r imputing missing values, echo=TRUE}
#computing steps mean by interval
activity2 <- activity %>%
  group_by(interval) %>%
  mutate(steps=ifelse(is.na(steps), mean(steps, na.rm=T), steps)) %>%
  ungroup()
```

#### Histogram of the total number of steps taken each day after missing values are imputed
```{r histogram with imputed missing values, echo=TRUE}
#plotting the new dataset
activity2 %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarise(total_steps=sum(steps, na.rm=T)) %>%
  ungroup() %>%
  ggplot(aes(x = total_steps)) +
        geom_histogram(binwidth = 1000) +
        labs(title = "Daily Steps", x = "Total Steps", y = "Frequency")
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r weekdays, echo=TRUE}
activity2 %>%
  mutate(weektime=ifelse(weekdays(date) %in% c("Samedi", "Dimanche"), "weekend", "weekday")) %>%
  group_by(interval, weektime) %>%
  summarise(mean=mean(steps, na.rm=T)) %>%
  ungroup() %>%
  ggplot(aes(x=interval, y=mean, color=weektime)) +
  geom_line() +
  facet_grid(weektime~.)
```

